---
title: "Duke Offensive Stats: 2022-23"
format: pdf
editor: visual
---

## Packages

```{r}
#| label: load-packages

library(tidyverse)
library(tidymodels)
```

# Home-Game Attendance & Offensive Performance

This section explores if any relationship appears to exist between game attendance and the offensive performance of Duke during games in Wallace Wade Stadium.

## Import Data

```{r}
#| label: load-data

offense_data <- read_csv("data/Duke Stats - DukeOffense.csv")

offense_data <- offense_data |>
  mutate(isHome = if_else(Site == "Home", TRUE, FALSE)) |>
  mutate(Day = as.factor(Day)) |>
  mutate(AttPct = if_else(AttNum/40004 > 1.0, 100.0, AttNum/40004*100))

home_offense_data <- offense_data |>
  filter(isHome == TRUE)

glimpse(home_offense_data)
```

Not all columns are used for each type of offensive statistic. For example, the "Onside" column is only relevant for rows whose *Type* column value is "Kickoffs". The *Comp* column represents completions (in terms of completed passes), successes (with Field_Goals, 3rd_Down_Conv, 4th_Down_Conv, etc.), or a total count (with Duke_Penalties, Opp_Penalties, etc.) depending on the football context of the row's *Type*.

## Rushing

#### Attendance as a predictor of *average* *yards* gained/lost per rushing play:

```{r}
#| label: attendance-vs-rushing-avg-yards

# Dataset filtering
home_off_rush_data <- home_offense_data |>
  filter(Type == "Rushing")

# Visualization
home_off_rush_data |>
  ggplot(
    aes(x = AttPct, y = AvgYd, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ log(x), se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. Average Yards per Rushing Play",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "Avg. Yards per Rush",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_rush_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(AvgYd ~ log(AttPct), data = home_off_rush_data)

tidy(att_rush_glm)
glance(att_rush_glm)$AIC

```

Wallace Wade attendance was *not* a strongly significant predictor of average yards gained/lost per rushing play in 2022-23.

## Passing

#### Attendance as a predictor of *average* *yards* gained/lost per passing play:

```{r}
#| label: attendance-vs-passing-avg-yards

# Dataset filtering
home_off_pass_data <- home_offense_data |>
  filter(Type == "Passing")

# Visualization
home_off_pass_data |>
  ggplot(
    aes(x = AttPct, y = AvgYd, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ x, se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. Average Yards per Passing Play",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "Avg. Yards per Pass",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_pass_yd_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(AvgYd ~ AttPct, data = home_off_pass_data)

tidy(att_pass_yd_glm)
glance(att_pass_yd_glm)$AIC
```

Wallace Wade attendance was *not* a statistically significant predictor of average yards gained/lost per passing play in 2022-23.

#### Attendance as a predictor of passing *completions* per game:

```{r}
#| label: attendance-vs-passing-completions

# Visualization
home_off_pass_data |>
  ggplot(
    aes(x = AttPct, y = CompPct, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ log(x), se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  scale_y_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. Pass Completions",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "% of Passing Attempts Ruled Complete",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_pass_comp_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(CompPct ~ log(AttPct), data = home_off_pass_data)

tidy(att_pass_comp_glm)
glance(att_pass_comp_glm)$AIC
```

Wallace Wade attendance was *not* a statistically significant predictor of the percentage of passing plays that were completed per game in 2022-23.

#### Attendance as a predictor of *touchdown* passes per game:

```{r}
#| label: attendance-vs-passing-touchdowns

# Visualization
home_off_pass_data |>
  ggplot(
    aes(x = AttPct, y = TD_Gained, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ x, se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. Touchdown Passes",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "Number of Passing Touchdowns",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_pass_td_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(TD_Gained ~ AttPct, data = home_off_pass_data)

tidy(att_pass_td_glm)
glance(att_pass_td_glm)$AIC
```

Wallace Wade attendance was *not* a statistically significant predictor of the number of touchdown passing plays per game in 2022-23.

#### Attendance as a predictor of *pass rating*:

```{r}
#| label: attendance-vs-passing-rating

# Visualization
home_off_pass_data |>
  ggplot(
    aes(x = AttPct, y = Rating, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ x, se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. Passing Rating",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "Pass Rating (higher is better)",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_pass_qb_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(Rating ~ AttPct, data = home_off_pass_data)

tidy(att_pass_qb_glm)
glance(att_pass_qb_glm)$AIC
```

Wallace Wade attendance was *not* a statistically significant predictor of passing rating per game in 2022-23.

# Punt Returns

#### Attendance as a predictor of *average* *yards* returned per punt return:

```{r}
#| label: attendance-vs-punt-returns-avg-yards

# Dataset filtering
home_off_punt_return_data <- home_offense_data |>
  filter(Type == "Punt_Returns")

# Visualization
home_off_punt_return_data |>
  ggplot(
    aes(x = AttPct, y = AvgYd, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ log(x), se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. Average Yards per Punt Return",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "Avg. Yards per Punt Return (higher is better)",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_punt_ret_yd_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(AvgYd ~ log(AttPct), data = home_off_punt_return_data)

tidy(att_punt_ret_yd_glm)
glance(att_punt_ret_yd_glm)$AIC
```

Wallace Wade attendance was *not* a statistically significant predictor of average punt return yardage per game in 2022-23.

#### Attendance as a predictor of punt return *attempts* per game:

```{r}
#| label: attendance-vs-punt-returns-attempts

# Visualization
home_off_punt_return_data |>
  ggplot(
    aes(x = AttPct, y = Attempts, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ log(x), se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. Punt Return Attempts",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "Number of Punt Returns",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_punt_ret_attempts_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(Attempts ~ log(AttPct), data = home_off_punt_return_data)

tidy(att_punt_ret_attempts_glm)
glance(att_punt_ret_attempts_glm)$AIC
```

Wallace Wade attendance was *not* a statistically significant predictor of the number of punt return attempts per game in 2022-23.

# Kickoff Returns

#### Attendance as a predictor of *average* *yards* returned per kickoff return:

```{r}
#| label: attendance-vs-kickoff-returns-avg-yards

# Dataset filtering
home_off_kickoff_return_data <- home_offense_data |>
  filter(Type == "Kickoff_Returns")

# Visualization
home_off_kickoff_return_data |>
  ggplot(
    aes(x = AttPct, y = AvgYd, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ x, se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. Average Yards per Kickoff Return",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "Avg. Yards per Kickoff Return (higher is better)",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_kickoff_ret_yd_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(AvgYd ~ AttPct, data = home_off_kickoff_return_data)

tidy(att_kickoff_ret_yd_glm)
glance(att_kickoff_ret_yd_glm)$AIC
```

Wallace Wade attendance was *not* a statistically significant predictor of the average yards returned per kickoff return attempt in 2022-23.

#### Attendance as a predictor of kickoff return *attempts* per game:

```{r}
#| label: attendance-vs-kickoff-returns-attempts

# Visualization
home_off_kickoff_return_data |>
  ggplot(
    aes(x = AttPct, y = Attempts, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ log(x), se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. Kickoff Return Attempts",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "Number of Kickoff Returns",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_kickoff_ret_attempts_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(Attempts ~ log(AttPct), data = home_off_kickoff_return_data)

tidy(att_kickoff_ret_attempts_glm)
glance(att_kickoff_ret_attempts_glm)$AIC
```

Wallace Wade attendance was *not* a statistically significant predictor of the number of kickoff return attempts per game in 2022-23.

# Punts

#### Attendance as a predictor of *average* *yards* per punt:

```{r}
#| label: attendance-vs-punts-avg-yards

# Dataset filtering
home_off_punts_data <- home_offense_data |>
  filter(Type == "Punts")

# Visualization
home_off_punts_data |>
  ggplot(
    aes(x = AttPct, y = AvgYd, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ exp(x), se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. Average Yards per Punt",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "Avg. Yards per Punt",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_punts_yd_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(AvgYd ~ exp(AttPct), data = home_off_punts_data)

tidy(att_punts_yd_glm)
glance(att_punts_yd_glm)$AIC
```

Wallace Wade attendance was *not* a strong predictor of the average yards punted per punt attempt in 2022-23.

#### Attendance as a predictor of punting *attempts* per game:

```{r}
#| label: attendance-vs-punts-attempts

# Visualization
home_off_punts_data |>
  ggplot(
    aes(x = AttPct, y = Attempts, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ log(x), se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. Punting Attempts",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "Number of Punts",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_punts_attempts_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(Attempts ~ log(AttPct), data = home_off_punts_data)

tidy(att_punts_attempts_glm)
glance(att_punts_attempts_glm)$AIC
```

Wallace Wade attendance was *not* a statistically significant predictor of the number of punting attempts per game in 2022-23.

# Kickoffs

#### Attendance as a predictor of *average* *yards* per kickoff:

```{r}
#| label: attendance-vs-kickoffs-avg-yards

# Dataset filtering
home_off_kickoffs_data <- home_offense_data |>
  filter(Type == "Kickoffs")

# Visualization
home_off_kickoffs_data |>
  ggplot(
    aes(x = AttPct, y = AvgYd, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ poly(x,2), se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. Average Yards per Kickoff",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "Avg. Yards per Kickoff",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_kickoffs_yd_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(AvgYd ~ poly(AttPct,2), data = home_off_kickoffs_data)

tidy(att_kickoffs_yd_glm)
glance(att_kickoffs_yd_glm)$AIC
```

Wallace Wade attendance was *not* a strong predictor of the average yards kicked per kickoff attempt in 2022-23.

#### Attendance as a predictor of *touchback percentage* per game:

```{r}
#| label: attendance-vs-kickoffs-touchback-pct

# Visualization
home_off_kickoffs_data |>
  ggplot(
    aes(x = AttPct, y = TouchbackPct, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ poly(x,2), se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  scale_y_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. Touchbacks per Game",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "% Touchbacks of all Kickoffs",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_kickoffs_touchback_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(TouchbackPct ~ poly(AttPct,2), data = home_off_kickoffs_data)

tidy(att_kickoffs_touchback_glm)
glance(att_kickoffs_touchback_glm)$AIC
```

Wallace Wade attendance is a statistically ***possible*** predictor of the percent of kickoff attempts that were touchbacks in 2022-23.

# Field Goals

#### Attendance as a predictor of 3-point field goal percentage per game:

```{r}
#| label: attendance-vs-field-goal-pct

# Dataset filtering
home_off_fg_data <- home_offense_data |>
  filter(Type == "Field_Goals")

# Visualization
home_off_fg_data |>
  ggplot(
    aes(x = AttPct, y = CompPct, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ exp(x), se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  scale_y_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. 3-Pt. Field Goal Percentage",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "% of Field Goal Attempts that are Successful",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_fg_comp_pct_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(CompPct ~ exp(AttPct), data = home_off_fg_data)

tidy(att_fg_comp_pct_glm)
glance(att_fg_comp_pct_glm)$AIC
```

Based on Duke football games in Wallace Wade during 2022-23, there ***is*** significant evidence that when football attendance is 100% or very close to 100%, the predicted 3-pt. field goals percentage decreases dramatically.

#### Attendance as a predictor of the number of 3-pt. field goals *attempted* per game:

```{r}
#| label: attendance-vs-fg-attempts

# Visualization
home_off_fg_data |>
  ggplot(
    aes(x = AttPct, y = Attempts, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ x, se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. 3-Pt. Field Goal Attempts",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "Number of Field Goal Attempts",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_fg_attempts_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(Attempts ~ AttPct, data = home_off_fg_data)

tidy(att_fg_attempts_glm)
glance(att_fg_attempts_glm)$AIC
```

Wallace Wade attendance was *not* a statistically significant predictor of the number of 3-point field goals attempted per game in 2022-23.

#### Attendance as a predictor of the number of 3-pt. field goals *made* per game:

```{r}
#| label: attendance-vs-field-goal-comp-count
#| 
# Visualization
home_off_fg_data |>
  ggplot(
    aes(x = AttPct, y = Comp, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ exp(x), se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. 3-Pt. Field Goals Made",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "Number of Field Goals Made",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_fg_comp_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(Comp ~ exp(AttPct), data = home_off_fg_data)

tidy(att_fg_comp_glm)
glance(att_fg_comp_glm)$AIC
```

Wallace Wade attendance was *not* a *strong* predictor of the number of 3-point field goals made per game in 2022-23. (However, there may be a slightly stronger, albiet still weak, association between FGs made & stadium attendance than the association between FGs attempted & stadium attendance.)

# Total Offense

#### Attendance as a predictor of *average* *yards* per offensive play:

```{r}
#| label: attendance-vs-total-avg-yards

# Dataset filtering
home_off_total_data <- home_offense_data |>
  filter(Type == "Total_Offense")

# Visualization
home_off_total_data |>
  ggplot(
    aes(x = AttPct, y = AvgYd, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ exp(x), se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. Average Yards per Offensive Play",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "Avg. Yards per Play",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_total_yd_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(AvgYd ~ exp(AttPct), data = home_off_total_data)

tidy(att_total_yd_glm)
glance(att_total_yd_glm)$AIC
```

Wallace Wade attendance was *not* a statistically significant predictor of the average yards gained/lost per offensive play attempt in 2022-23.

#### Attendance as a predictor of total offensive *attempts* per game:

```{r}
#| label: attendance-vs-total-attempts

# Visualization
home_off_total_data |>
  ggplot(
    aes(x = AttPct, y = Attempts, color = FPI_diff)
  ) +
  geom_point() +
  geom_smooth(method = "glm", formula = y ~ poly(x,2), se = FALSE) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(title = "Stadium Attendance vs. Offensive Attempts",
       subtitle = "Duke Home-Field Football Games, 2022-23",
       x = "Percentage of Stadium Capacity Filled",
       y = "Number of Offensive Attempts",
       color = "FPI Difference\n(Higher = More\nDifficult Opponent)")

# Linear model
att_total_attempts_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(Attempts ~ poly(AttPct,2), data = home_off_total_data)

tidy(att_total_attempts_glm)
glance(att_total_attempts_glm)$AIC
```

Wallace Wade attendance was *not* a statistically significant predictor of the number of offensive attempts per game in 2022-23.
